Semantic Linearity in a Lazy Optimising Compiler

### 2024-02-12 icfp.tex
proof of commuting case-let

  {\small
  \[
    \begin{array}{ll}
   (\llet{\xD = e}{e'} )  \Longrightarrow \llet{\xD =       e}{e'[e/x]} &                    \mbox{Inlining}\\
       (\lambda \x[\pi][\s].~e)~e'  \Longrightarrow e[e'/x]  &                \mbox{$\beta$-reduction}\\
      (\ccase{(K~\ov{e})}{\z[\D][\s]~\{\overline{\rho_j\to e_j}, K~\ov{x} \to e_i, \overline{\rho_k\to e_k}\}} )\Longrightarrow  e_i\ov{[e/x]}[K~\ov{e}/z] &
                                  \mbox{Case of known constr.}\\
      (\ccase{(\ccase{e_c}{\z[\D]~\{\ov{\rho_{c_i}\to e_{c_i}}\}})}{\var[w][\irr{\D,\D'}][\s']~\{\ov{\rho_i\to e_i}\}})\\
            \Longrightarrow 
      \ccase{e_c}{\z[\D]~\{\ov{\rho_{c_i} \to \ccase{e_{c_i}}{w~\{\ov{\rho_i\to e_i}\}}}\}}
      & \mbox{Case-of-case}\\
      (\lambda \y[\pi].~\llet{\xD = e}{e'})
      \Longrightarrow
      \llet{\xD=e}{\lambda\y[\pi].~e'} & \mbox{Commuting let-$\lambda$}\\
      ((\llet{v = e}{b})~a) \Longrightarrow \llet{v = e}{b~a} &\mbox{Commuting let-app.}\\
      (\ccase{(\llet{v = e}{b})}{\{\overline{\rho\to e'}\}}) \Longrightarrow
      \llet{v =  e}{\ccase{b}{\{\overline{\rho\to e'}\}}}
                               &
                                 \mbox{Commuting let-case}\\
      (\ccase{e}{\{\overline{\rho_j\to e_j}, \rho \to E[e_1], \overline{\rho_k\to e_k}\}} ) \qquad (\mbox{with $x$ fresh}) \\
     \Longrightarrow \llet{x = e_1}{\ccase{e}{\{\overline{\rho_j\to e_j}, \rho \to E[t], \overline{\rho_k\to e_k}\}}} 
                                                                        & \mbox{Commuting case-let}\\
      (\llet{x = (\llet{v = e}{b})}{c}) \Longrightarrow \llet{v = e}{\llet{x = b}{c}}
                                                                      & \mbox{Commuting let-let}\\
      f \Longrightarrow \lambda x.(f~x) & \mbox{$\eta$-expansion}\\
      \lambda x.(f~x) \Longrightarrow f & \mbox{$\eta$-reduction}\\
      (\ccase{x}{z~\{\ov{\rho_i\to e_i}\}}) \Longrightarrow \ccase{x}{z~\{\ov{\rho_i\to  e_i[z/x]}\}}
                                                                      & \mbox{Binder swap}\\
      % (\ccase{x}{z~\{\ov{\rho_i\to e_i}\}} )\Longrightarrow \ccase{x}{z~\{\ov{\rho_i\to e_i[x/z]}\}}
      %                                                                 & \mbox{Reverse binder swap}
    \end{array}
  \]}
  \caption{Optimising transformations validated by Linear Core\label{fig:opttrans} (see Appendix~\ref{app:optimisations} for details)} 
\end{figure}

>>>>

  {\small
  \[
    \begin{array}{ll}
   (\llet{\xD = e}{e'} )  \Longrightarrow \llet{\xD =       e}{e'[e/x]} &                    \mbox{Inlining}\\
       (\lambda \x[\pi][\s].~e)~e'  \Longrightarrow e[e'/x]  &                \mbox{$\beta$-reduction}\\
      (\ccase{(K~\ov{e})}{\z[\D][\s]~\{\overline{\rho_j\to e_j}, K~\ov{x} \to e_i, \overline{\rho_k\to e_k}\}} )\Longrightarrow  e_i\ov{[e/x]}[K~\ov{e}/z] &
                                  \mbox{Case of known constr.}\\
      (\ccase{(\ccase{e_c}{\z[\D]~\{\ov{\rho_{c_i}\to e_{c_i}}\}})}{\var[w][\irr{\D,\D'}][\s']~\{\ov{\rho_i\to e_i}\}})\\
            \Longrightarrow 
      \ccase{e_c}{\z[\D]~\{\ov{\rho_{c_i} \to \ccase{e_{c_i}}{w~\{\ov{\rho_i\to e_i}\}}}\}}
      & \mbox{Case-of-case}\\
      (\lambda \y[\pi].~\llet{\xD = e}{e'})
      \Longrightarrow
      \llet{\xD=e}{\lambda\y[\pi].~e'} & \mbox{Commuting let-$\lambda$}\\
      ((\llet{v = e}{b})~a) \Longrightarrow \llet{v = e}{b~a} &\mbox{Commuting let-app.}\\
      (\ccase{(\llet{v = e}{b})}{\{\overline{\rho\to e'}\}}) \Longrightarrow
      \llet{v =  e}{\ccase{b}{\{\overline{\rho\to e'}\}}}
                               &
                                 \mbox{Commuting let-case}\\
      (\ccase{e}{\{\overline{\rho_j\to e_j}, \rho \to E[e_1],
      \overline{\rho_k\to e_k}\}} ) \qquad (\mbox{$x$ fresh}) \\
     \Longrightarrow \llet{x = e_1}{\ccase{e}{\{\overline{\rho_j\to e_j}, \rho \to E[x], \overline{\rho_k\to e_k}\}}} 
                                                                        & \mbox{Commuting case-let}\\
      (\llet{x = (\llet{v = e}{b})}{c}) \Longrightarrow \llet{v = e}{\llet{x = b}{c}}
                                                                      & \mbox{Commuting let-let}\\
      f \Longrightarrow \lambda x.(f~x) & \mbox{$\eta$-expansion}\\
      \lambda x.(f~x) \Longrightarrow f & \mbox{$\eta$-reduction}\\
      (\ccase{x}{z~\{\ov{\rho_i\to e_i}\}}) \Longrightarrow \ccase{x}{z~\{\ov{\rho_i\to  e_i[z/x]}\}}
                                                                      & \mbox{Binder swap}\\
      % (\ccase{x}{z~\{\ov{\rho_i\to e_i}\}} )\Longrightarrow \ccase{x}{z~\{\ov{\rho_i\to e_i[x/z]}\}}
      %                                                                 & \mbox{Reverse binder swap}
    \end{array}
  \]}
  \caption{Optimising transformations validated by Linear
    Core\label{fig:opttrans} (for commuting case-let we
      require $e_1$ not to capture the case binder, variables bound by the
      pattern $\rho$ and the context $E[{-}]$ -- see Appendix~\ref{app:optimisations} for
    details).} 
\end{figure}
